---
title: "R Notebook"
output: html_notebook
---



Instructions
100 XP
Filter for only the year 1990.
Sort the table in descending order of the number of babies born.

```{r}

install.packages("babynames")
library(babynames)
head(babynames)
babynames <- readRDS("babynames.rds")
```

```{r}
library(dplyr)
babynames %>% filter(year == 1990) %>% arrange(desc(n))
  # Filter for the year 1990
  
  # Sort the number column in descending order 
 
```

Exercise
Exercise
Using top_n with babynames
You saw that you could use filter() and arrange() to find the most common names in one year. However, you could also use group_by and top_n to find the most common name in every year.



```{r}
# Find the most common name in each year
babynames %>% group_by(year) %>% top_n(1, number)
  
  
```
Visualizing names with ggplot2
The dplyr package is very useful for exploring data, but it's especially useful when combined with other tidyverse packages like ggplot2.

Instructions 1/2
50 XP
Filter for only the names Steven, Thomas, and Matthew, and assign it to an object called selected_names.


```{r}
# Filter for the names Steven, Thomas, and Matthew 
selected_names <- babynames %>% filter(name %in% c("Steven","Thomas","Matthew"))
selected_names

# Plot the names using a different color for each name
ggplot(selected_names, aes(x = year, y = number, color = name)) +
  geom_line()


```
Finding the year each name is most common
In an earlier video, you learned how to filter for a particular name to determine the frequency of that name over time. Now, you're going to explore which year each name was the most common.

To do this, you'll be combining the grouped mutate approach with a top_n.

Instructions 1/2
0 XP
First, calculate the total number of people born in that year in this dataset as year_total.
Next, use year_total to calculate the fraction of people born in each year that have each name.



```{r}

# Calculate the fraction of people born each year with the same name
babynames_fraction <- babynames %>%
  group_by(year) %>%
  mutate(year_total = sum(number)) %>%
  ungroup() %>%
  mutate(fraction = number / year_total) %>%
# Find the year each name is most common
  group_by(name) %>%
  top_n(1,fraction)
babynames_fraction
```

Adding the total and maximum for each name
In the video, you learned how you could group by the year and use mutate() to add a total for that year.

In these exercises, you'll learn to normalize by a different, but also interesting metric: you'll divide each name by the maximum for that name. This means that every name will peak at 1.

Once you add new columns, the result will still be grouped by name. This splits it into 48,000 groups, which actually makes later steps like mutates slower.

Instructions 1/2
50 XP
Instructions 1/2
50 XP
Use a grouped mutate to add two columns:
name_total, with the total number of babies born with that name in the entire dataset.
name_max, with the highest number of babies born in any year.


Add another step to ungroup the table.
Add a column called fraction_max, with the number in the year divided by the maximum for that name.




```{r}

# Add columns name_total and name_max for each name
babynames %>% group_by(name) %>% mutate(name_total = sum(number), name_max = max(number))  %>%
  # Ungroup the table 
  ungroup() %>% mutate(fraction_max = number/ name_max)
  # Add the fraction_max column containing the number by the name maximum 
  

```
Visualizing the normalized change in popularity
You picked a few names and calculated each of them as a fraction of their peak. This is a type of "normalizing" a name, where you're focused on the relative change within each name rather than the overall popularity of the name.

In this exercise, you'll visualize the normalized popularity of each name. Your work from the previous exercise, names_normalized, has been provided for you.
```{r}

names_normalized <- babynames %>%
                     group_by(name) %>%
                     mutate(name_total = sum(number),
                            name_max = max(number)) %>%
                     ungroup() %>%
                     mutate(fraction_max = number / name_max)
```

Instructions
100 XP
Instructions
100 XP
Filter the names_normalized table to limit it to the three names Steven, Thomas, and Matthew.
Visualize fraction_max for those names over time.


```{r}

# Filter for the names Steven, Thomas, and Matthew
names_filtered <- names_normalized %>%
  filter(name %in% c("Steven", "Thomas", "Matthew"))

# Visualize these names over time
ggplot(names_filtered, aes(x = year, y = fraction_max, color = name)) +
  geom_line()
```

Using ratios to describe the frequency of a name
In the video, you learned how to find the difference in the frequency of a baby name between consecutive years. What if instead of finding the difference, you wanted to find the ratio?

You'll start with the babynames_fraction data already, so that you can consider the popularity of each name within each year.

Instructions
100 XP
Arrange the data in ascending order of name and then year.
Group by name so that your mutate works within each name.
Add a column ratio containing the ratio (not difference) between each year.

Hint
Similar to the difference column from the video, the ratio column will also be calculated using lag(fraction).
Recall, the difference in the frequency of a baby name between consecutive years was calculated as fraction - lag(fraction). However, remember that in this exercise you are calculating the ratio, so another operator will be required.

```{r}
babynames_fraction %>%
  # Arrange the data in order of name, then year 
  arrange(name, year) %>%
  # Group the data by name
  group_by(name) %>%
  # Add a ratio column that contains the ratio between each year 
  mutate(ratio = fraction / lag(fraction))

```

Biggest jumps in a name
Previously, you added a ratio column to describe the ratio of the frequency of a baby name between consecutive years to describe the changes in the popularity of a name. Now, you'll look at a subset of that data, called babynames_ratios_filtered, to look further into the names that experienced the biggest jumps in popularity in consecutive years.
```{r}

babynames_ratios_filtered <- babynames_fraction %>%
                     arrange(name, year) %>%
                     group_by(name) %>%
                     mutate(ratio = fraction / lag(fraction)) %>%
                     filter(fraction >= 0.00001)

```

babynames_ratios_filtered <- babynames_fraction %>%
                     arrange(name, year) %>%
                     group_by(name) %>%
                     mutate(ratio = fraction / lag(fraction)) %>%
                     filter(fraction >= 0.00001)
Instructions
100 XP
Instructions
100 XP
From each name in the data, keep the observation (the year) with the largest ratio; note the data is already grouped by name.
Sort the ratio column in descending order.
Filter the babynames_ratios_filtered data further by filtering the fraction column to only display results greater than or equal to 0.001.
```{r}
babynames_ratios_filtered %>%
  # Extract the largest ratio from each name 
  top_n(1,ratio) %>%
  # Sort the ratio column in descending order 
  arrange(desc(ratio)) %>%
  # Filter for fractions greater than or equal to 0.001
filter(fraction >= 0.001)



```
