---
title: "R Notebook"
output: html_notebook
---

Selecting columns
Using the select verb, we can answer interesting questions about our dataset by focusing in on related groups of verbs. The colon (:) is useful for getting many columns at a time.

Instructions
100 XP
Instructions
100 XP
Use glimpse() to examine all the variables in the counties table.
Select the columns for state, county, population, and (using a colon) all five of those industry-related variables; there are five consecutive variables in the table related to the industry of people's work: professional, service, office, construction, and production.
Arrange the table in descending order of service to find which counties have the highest rates of working in the service industry.

```{r}

counties <- readRDS("counties.rds")

head(counties)

```

Selecting columns
Using the select verb, we can answer interesting questions about our dataset by focusing in on related groups of verbs. The colon (:) is useful for getting many columns at a time.

Instructions
100 XP
Instructions
100 XP
Use glimpse() to examine all the variables in the counties table.
Select the columns for state, county, population, and (using a colon) all five of those industry-related variables; there are five consecutive variables in the table related to the industry of people's work: professional, service, office, construction, and production.
Arrange the table in descending order of service to find which counties have the highest rates of working in the service industry.


```{r}
# Glimpse the counties table
glimpse(counties)
colnames(counties)
counties %>% select(state,county,population, ("professional":"production")) %>% arrange(desc(service))
  # Select state, county, population, and industry-related columns
  
  # Arrange service in descending order 

```
Select helpers
In the video you learned about the select helper starts_with(). Another select helper is ends_with(), which finds the columns that end with a particular string.

Instructions
100 XP
Select the columns state, county, population, and all those that end with work.
Filter just for the counties where at least 50% of the population is engaged in public work.

```{r}
counties %>% select(state,county,population,ends_with("work")) %>% filter(public_work >= 50)
  # Select the state, county, population, and those ending with "work"
 
  # Filter for counties that have at least 50% of people engaged in public work



```

Renaming a column after count
The rename() verb is often useful for changing the name of a column that comes out of another verb, such as count(). In this exercise, you'll rename the n column from count() (which you learned about in Chapter 2) to something more descriptive.

Instructions 1/2
50 XP
Instructions 1/2
50 XP
Use count() to determine how many counties are in each state.

Notice the n column in the output; use rename() to rename that to num_counties.

```{r}

# Rename the n column to num_counties
counties %>%
  count(state) %>% rename(num_counties = n)

```

Renaming a column as part of a select
rename() isn't the only way you can choose a new name for a column: you can also choose a name as part of a select().

Instructions
100 XP
Select the columns state, county, and poverty from the counties dataset; in the same step, rename the poverty column to poverty_rate.

```{r}

# Select state, county, and poverty as poverty_rate
counties %>% select(state,county, poverty_rate = poverty)
 

```

Using transmute
As you learned in the video, the transmute verb allows you to control which variables you keep, which variables you calculate, and which variables you drop.

Instructions
100 XP
Keep only the state, county, and population columns, and add a new column, density, that contains the population per land_area.
Filter for only counties with a population greater than one million.
Sort the table in ascending order of density.


```{r}

counties %>%
  # Keep the state, county, and populations columns, and add a density column
  transmute(state,county,population,density = population/land_area) %>%
  # Filter for counties with a population greater than one million 
  filter(population > 1000000) %>% arrange(density)
  # Sort density in ascending order 
  

```
Choosing among the four verbs
In this chapter you've learned about the four verbs: select, mutate, transmute, and rename. Here, you'll choose the appropriate verb for each situation. You won't need to change anything inside the parentheses.

Instructions
100 XP
Instructions
100 XP
Choose the right verb for changing the name of the unemployment column to unemployment_rate
Choose the right verb for keeping only the columns state, county, and the ones containing poverty.
Calculate a new column called fraction_women with the fraction of the population made up of women, without dropping any columns.
Keep only three columns: the state, county, and employed / population, which you'll call employment_rate.

```{r}
# Change the name of the unemployment column
counties %>%
  rename(unemployment_rate = unemployment)

# Keep the state and county columns, and the columns containing poverty
counties %>%
  select(state, county, contains("poverty"))

# Calculate the fraction_women column without dropping the other columns
counties %>%
  mutate(fraction_women = women / population)

# Keep only the state, county, and employment_rate columns
counties %>%
  transmute(state, county, employment_rate = employed / population)
```

